Urban Crime Dynamics: Trends in Theft and Assault Over 2014 to 2023

### 2024-09-26 paper/paper.qmd
change the graph 1 code and the name of graph 1

## Graph 1
\
```{r}
#| echo: false
#| warning: false
#### First Graph ####

library(ggplot2)
library(reshape2)
# This code is based on sample from chatGPT

# Select the columns related to theft rates (2014-2023)
theft_data <- crime_rates_data[, c("AREA_NAME", "THEFTOVER_RATE_2014", "THEFTOVER_RATE_2015", 
                             "THEFTOVER_RATE_2016", "THEFTOVER_RATE_2017", 
                             "THEFTOVER_RATE_2018", "THEFTOVER_RATE_2019",
                             "THEFTOVER_RATE_2020", "THEFTOVER_RATE_2021", 
                             "THEFTOVER_RATE_2022", "THEFTOVER_RATE_2023")]

theft_data_melted <- melt(theft_data, id.vars = "AREA_NAME", variable.name = "Year", value.name = "TheftRate")

# Clean up the contents of the Year column by removing the prefix 'THEFTOVER_RATE_'
theft_data_melted$Year <- gsub("THEFTOVER_RATE_", "", theft_data_melted$Year)
```
\

```{r}
#| echo: false
#| warning: false
#| label: fig-theft-rate-trend-2014-to-2023
#| fig-cap: "Trends in theft rates by region from 2014 to 2023"

# This code is based on sample from chatGPT

# Plot the trend of theft rates in each region over the years
ggplot(theft_data_melted, aes(x = Year, y = TheftRate, group = AREA_NAME, color = AREA_NAME)) +
  geom_line() +
  labs(title = "Trends in theft rates (2014-2023)",
       x = "Year",
       y = "Theft rate (per 100,000 people)") +
  theme_minimal() +
  theme(legend.position = "none")
```

>>>>

## Graph 1

```{r}
#| label: fig-theft-rate-trend-2014-to-2023
#| echo: false
#| warning: false
#| fig-cap: "Trends in theft Rates in the 15 neighborhoods with the Highest Average theft Rates and the 15 neighborhoods with the Lowest Theft Rates, 2014-2023"

#### First Graph ####
library(tidyverse)
# This code is based on sample from ChatGPT

# Select relevant columns and gather year and theft rate into long format
crime_rates_long <- crime_rates_data %>%
  select(AREA_NAME, starts_with("THEFTOVER_RATE")) %>%
  gather(key = "Year", value = "TheftRate", -AREA_NAME) %>%
  mutate(Year = as.integer(sub("THEFTOVER_RATE_", "", Year)))

# Calculate average theft rate across the years for each city
average_theft_rates <- crime_rates_long %>%
  group_by(AREA_NAME) %>%
  summarise(AverageTheftRate = mean(TheftRate, na.rm = TRUE)) %>%
  ungroup()

# Select the top 15 and bottom 15 cities based on average theft rates
top_cities <- top_n(average_theft_rates, 15, AverageTheftRate)
bottom_cities <- top_n(average_theft_rates, -15, AverageTheftRate)

# Filter the original long data for these cities
filtered_data <- crime_rates_long %>%
  filter(AREA_NAME %in% c(top_cities$AREA_NAME, bottom_cities$AREA_NAME))

# Create the plot
ggplot(filtered_data, aes(x = Year, y = TheftRate, color = AREA_NAME)) +
  geom_line() +
  labs(title = "Trends in Theft Rates (2014-2023)",
       subtitle = "Top 15 and Bottom 15 Cities by Average Theft Rate",
       x = "Year",
       y = "Theft rate per 100,000 people") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(nrow = 2, byrow = TRUE))
```
```
